Hedonic Price Modelling with Geographically Weighted Regression in Medellín, Colombia

### 2016-07-26 paper/paper_GWR.tex
Rendering.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/tex-archive/info/epslatex/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{Bare Demo of IEEEtran.cls for Journals}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Michael~Shell,~\IEEEmembership{Member,~IEEE,}
        John~Doe,~\IEEEmembership{Fellow,~OSA,}
        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
\thanks{M. Shell is with the Department
of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
\thanks{Manuscript received April 19, 2005; revised December 27, 2012.}}

@@

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~11, No.~4, December~2012}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
IEEEtran, journal, \LaTeX, paper, template.
\end{IEEEkeywords}

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{E}{stimating} the prices of real estate is useful and
necessary for taxation, public policy, infrastructure and
equality in cities. Medellín is the second city of
Colombia, with more than two and a half million
inhabitants and a population of over three million in its
metropolitan area. It is a focus of development for the
country, and construction plays a major role in its
dynamics.
Most of the development in Medellín is very uniform in
its types. The developers do not have a clear idea of what
customers want, and thus keep building over the same
formula that has worked this far. Universidad EAFIT
funded a project with its Research Group in Spatial
Economics (RiSE) to assess the real estate prices in
Medellín, in order to seek for what it is that costumers
value in housing.
Geographically Weighted Regression (GWR) was used
as a method for the hedonic pricing estimation of real
estate in Medellín using only structural variables (size,
number of rooms, parking, etc.) to predict total price.
Fair estimation results were obtained by using GWR, but
some questions arise:

@@

\item Hedonic pricing theory considers a group of
independent variables that can help explain the prices
of real estate and are related to the surroundings of a
housing unit such as urban form or density, number
of parks, accessibility, and quality of transportation,
among many others. What contextual variables can be meaningful to improve the actual estimation of real
estate prices in Medellín?
\item Can clustering techniques reveal housing
submarkets? Would they be similar as the one
experience has told developers in Medellín so far?
\item Does the geographical location of a housing unit
absorb other important variables in hedonic pricing
theory? How does it relate to them? Which can better
explain price?

@@

\textbf{LITERATURE REVIEW}

@@

The lengthiest part of this project was to process
geographical data from different sources to add to
the original data base that was used and analyzed
with Universidad EAFIT.

\begin{enumerate}
\item Housing data was obtained from OIME, Medellín’s
municipal Real Estate Market Observatory. It can be
found online 1 . Initially, it has 2003 observations for
new and used houses and apartments across the city
of Medellín in the year of 2014 (See Figure 1.).

@@

The analyzed dataset comprises the following
variables. The summary data was calculated after
removing atypical values and observations with
missing data (absolute z-scores greater than 3):

@@

\begin{itemize}
\item Medellín’s department of Planning Zoning plan 2 (ArcGIS geodatabase). Includes road networks, metro and BRT stations locations, existing parks, among others. The original maps can be
seen in Anexus 1 and Anexus 2. A good amount of geographical processing was needed in order to create variables for nearest distance to features for each of the data points, and to extract
the features that were needed in the case of existing green parks and only the main roads in the network (Anexus 3).

@@

\textbf{TABLE}

@@

\begin{itemize}
\item Density and Distance to Downtown are relatively highly positively correlated. This is only a demonstration of the known fact that most of Medellín lives in its periphery.

@@

Since inference is complicated in GWR estimation, no assertions can be made, only that the expected added value to green areas according to location has the expected behavior that is indeed observed
in Figure 7: wealthier areas are more attracted to this feature, whereas poorer areas actually repel it because it might usually mean less access to services.

@@

@@

\appendices
\section{Proof of the First Zonklar Equation}
Appendix one text goes here.

@@

The authors would like to thank...

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

@@

@@

\begin{IEEEbiography}{Michael Shell}
Biography text here.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

@@

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}

>>>>

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{/home/saf537/Documents/GWR_hedonicP_MDE/Plots}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png,.jpg}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{/home/saf537/Documents/GWR_hedonicP_MDE/Plots}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps,.pdf,.jpeg,.png,.jpg}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/tex-archive/info/epslatex/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{Hedonic Price Modelling with Geographically Weighted Regression in Medellín, Colombia}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Sara~Arango,
        Juan~Carlos~Duque.% <-this % stops a space
\thanks{Sara Arango Center for Urban Science and Progress, \textit{CUSP}. 
New York University, United States.}% <-this % stops a space
\thanks{Group in Spatial Economics, \textit{RiSE} 
Universidad EAFIT, Medellín, Colombia}}

@@

% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~11, No.~4, December~2012}%
%{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
In this paper we use Geographically Weighted Regression to estimate Real Estate Prices with an hedonic pricing approach, accounting for selected contextual and structural variables for city appraisal records of the Medellín, Colombia in 2014, and claiming that this type of analysis can be equally efficient for emergent cities such as Medellín, regardless of the size of their informal markets.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Hedonic pricing, real estate, Geographically Weighted Regression.
\end{IEEEkeywords}

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{E}{stimating} the prices of real estate is useful and
necessary for taxation, public policy, infrastructure and equality in cities.

\textbf{Hedonic pricing.} Hedonic pricing is an approach that has been naturally linked to real estate as it welcomes the modelling of assets that are usually not valued by traditional pricing strategies, such as natural resources, accesibility, among others.

GWR is a spatial method developed by Fotheringham et al. \cite{fother} to account for the fact that independent variables might have different effects on a dependent variable depending on spatial location. For example, the effect of distance to subway stations in a city in real estate prices may depend on geographical factors that might not be evident from other independent variables.

\textbf{GWR and hedonic pricing.}

Medellín is the second city of Colombia, with more than two and a half million inhabitants and a population of over three million in its metropolitan area. It is a focus of development for the country, and construction plays a major role in its dynamics. \textbf{CITA SOBRE MEDELLIN SIENDO IMPORTANTE EN REAL ESTATE}.

Most of the formal development in Medellín is very uniform in its types. The developers do not have a clear idea of what customers want, and thus they keep building over the same formula that has worked this far. \textit{Universidad EAFIT} funded a project with its Research Group in Spatial Economics (RiSE) to assess the real estate prices in Medellín, in order to seek for what it is that costumers value in housing, and to set foundations for a spin-off with this prpose.

GWR was used as a method for the hedonic pricing estimation of real estate in Medellín using structural variables (those particular to the housing unit itself) and contextual variables (those related to neighbour quality) to predict total price. An important amount of geographical processing had to be done in order to collect, process and standardize this information, typically unavailable for the real estate business to use.

Some questions arose:

@@

\item What contextual variables can be meaningful to improve the actual estimation of real estate prices in Medellín? 
\item Does the geographical location of a housing unit absorb other important variables in hedonic pricing theory? How does it relate to them? Which can better explain the price?

@@

\section{Methodology}

@@

A first important step of this project was to process geographical data from different sources to add to the original data base that was used and analyzed with Universidad EAFIT.

\begin{enumerate}
\item Housing data was obtained from OIME, Medellín’s municipal Real Estate Market Observatory. It can be found online 1 . Initially, it has 2003 observations for new and used houses and apartments across the city of Medellín in the year of 2014 (See Figure \ref{fig_sim}.).

\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{/home/saf537/Documents/GWR_hedonicP_MDE/Plots/MDE_point}
\caption{Distribution of data points.}
\label{fig_sim}
\end{figure}

@@

The analysed dataset comprises the following
variables. The summary data was calculated after
removing atypical values and observations with
missing data (absolute z-scores greater than 3):

@@

\begin{itemize}
\item Medellín’s department of Planning Zoning plan 2 (ArcGIS geodatabase). Includes road networks, metro and BRT stations locations, existing parks, among others. A good amount of geographical processing was needed in order to create variables for nearest distance to features for each of the data points, and to extract
the features that were needed in the case of existing green parks and only the main roads in the network.

@@

May be this is what you are after:

\begin{table}[htb]
\caption{Summary of the newly created variables.}
\label{1234}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{c|c|c|c|c|c}
value & \textbf{Density} & \textbf{Open space} & \textbf{Transit} & \textbf{Downtown }& \textbf{Roads} \\
\hline
mean & 24630 & 0.001 & 0.0113 & 0.040 & 0.002\\
std & 14576 & 0.001 & 0.0077 & 0.019 & 0.002 \\
min & 0 & 0.000 & 0.0006 & 0.001 & 0.000 \\
25\% & 14453 & 0.000 & 0.0055 & 0.027 & 0.001 \\
50\% & 21218 & 0.001 & 0.0099 & 0.039 & 0.001 \\
75\% & 32645 & 0.001 & 0.0153 & 0.052 & 0.003 \\
max & 77837 & 0.04 & 41.2 & 0.085 & 0.013 \\

\end{tabular}
}
\end{table}

@@

\begin{itemize}
\item Density and Distance to Downtown are relatively highly positively correlated. This is only a demonstration of the known fact that most of Medellín lives in its periphery \ref{cor}.

\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{/home/saf537/Documents/GWR_hedonicP_MDE/Plots/cor}
\caption{Distribution of data points.}
\label{cor}
\end{figure}

@@

Since inference is complicated in GWR estimation, no assertions can be made, only that the expected added value to green areas according to location has the expected behaviour that is indeed observed
in Figure 7: wealthier areas are more attracted to this feature, whereas poorer areas actually repel it because it might usually mean less access to services.

@@

Not necessarily a novel approach, but it is a novel exercise in the context of emerging cities with potentially very obscure sub markets.

It might be worth just modelling subsets of the city.

GWR must be taken and understood with care since it is a very novel technique and validation is harder, still to be more rigorous.

What is the best way to standardize? what is the best way to treat variables?

Feature selection.

@@

\appendices
\section{Geographical inputs}

@@

The authors would like to thank the City of Medellín's Real Estate Observatory (OIME) in the head of Juan Pablo Barrero.

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{2}

@@

\bibitem{fother}
Fotheringham, A. Stewart., Chris Brundson, and Martin Charlton. \emph{Geographically Weighted Regression: The Analysis of Spatially Varying Relationships}. Chichester, West Sussex, England: Wiley, 2002. Print.

@@

@@
